---
title: "Clinical Trial in Soft Tissue Sarcoma"
output: html_document
---
## Get ready with the data
```{r setup, results = 'hide', message = FALSE}
# list of packages used
pkg <- c("tidyverse", "readxl", "arsenal")

# activate packages
lapply(pkg, library, character.only = TRUE)

# grab source data
baseline <- read_excel(
  "SARC021_DATA.xlsx",
  sheet = "Baseline information"
)
last_response <- read_excel(
  "SARC021_DATA.xlsx",
  sheet = "Last Tumor Respsonse"
)
best_response <- read_excel(
  "SARC021_DATA.xlsx",
  sheet = "Best Tumor Response"
)
adverse_event <- read_excel(
  "SARC021_DATA.xlsx",
  sheet = "Adverse Event"
)

```

### Clean and Pre-process the Data
```{r func}
# this function provides table-1 data for categorical variables
cat_compare <- function(x) {
  # input x is a character string of a column name
  comps <- baseline %>%
    # summarize by var and arm to get number of subjects
    group_by(ARMCD, get(x)) %>%
    summarize(n = n()) %>%
    # get percentage of each value for each arm
    mutate(percent = round(100 * n / sum(n), 2)) %>%
    # make a character variable in format for table
    mutate(stat = paste0(n, " (", percent, "%)")) %>%
    # drop the intermediate vars
    select(-n,-percent) %>%
    ungroup() %>%
    # spread into the right structure for the table
    spread(key = ARMCD, value = stat) %>%
    # rename columns, create blank P column
    transmute(
      Characteristics = `get(x)`,
      ArmA = A,
      ArmB = B,
      P = ""
    ) %>%
    # catch any NA cases resulting from an instance of a characteristic only
    #   present in one arm but not the other
    mutate(ArmA = if_else(!is.na(ArmA), ArmA, "0 (0%)"),
           ArmB = if_else(!is.na(ArmB), ArmB, "0 (0%)"))
  
  # get p-val for homogeneity of proportions
  if (x == "PSURGFL") {
    # every value in this variable is "Y" so we have nothing to test
    p_val <- ""
  } else {
    p_val <- baseline %>%
      select(!!x, ARMCD) %>%
      table() %>%
      matrix(ncol = 2) %>%
      fisher.test(workspace = 2e7) %>%
      .$p.value %>%
      round(3)
  }
  
  # make header row with trait name and p-val
  head_row <-
    tibble(
      Characteristics = paste("-", x),
      ArmA = "",
      ArmB = "",
      P = as.character(p_val)
    )
  
  # bind and output
  rbind(head_row, comps)
}

```

## Baseline Values Comparison

```{r age_plot}
ggplot(baseline, aes(AGE)) +
  geom_histogram(bins = 30, fill = "#F8766D") +
  facet_wrap(~ ARMCD) +
  labs(title = "Histograms of age by study arm",
         x = "age")

```

The histograms of age for each study arm indicate that the distributions of ages are not normal. Thus, it is most appropriate to use a rank-sum test to compare the median ages between the two arms. This is how the p-value for the age composition across arms was obtained.

### Table. Baseline Information
``` {r q1}
# compute statistics for the age data
ages <- baseline %>%
  select(AGE, ARMCD) %>%
  group_by(ARMCD) %>%
  # all vars from example table
  summarize(
    mean = mean(AGE),
    n = n(),
    sd = sd(AGE),
    median = median(AGE),
    min = min(AGE),
    max = max(AGE)
  ) %>%
  select(-ARMCD) %>%
  signif(3) %>%
  # add back in arm info
  mutate(ARMCD = c("A", "B")) %>%
  # shift arm to first column
  select(ARMCD, everything())

# calculate p-value for age data
p_val <- kruskal.test(AGE ~ ARMCD, data = baseline) %>%
  .$p.value %>%
  signif(3)

# labels for table
Characteristics <- c("- AGE",
                     "Mean +- SD",
                     "Median (min, max)")

# combine results into strings for table
meanA <- paste0(ages[1, 2], " +- ", ages[1, 4])
medianA <-
  paste0(ages[1, 5], " (", ages[1, 6], ", ", ages[1, 7], ")")
ArmA <- c("",
          meanA,
          medianA)
meanB <- paste0(ages[2, 2], " +- ", ages[2, 4])
medianB <-
  paste0(ages[1, 5], " (", ages[1, 6], ", ", ages[1, 7], ")")
ArmB <- c("",
          meanB,
          medianB)
P <- c(p_val, "", "")

# table with ages data in the right structure for the table
age_tibble <- tibble(Characteristics, ArmA, ArmB, P)


# get the names of all the character columns needing analysis
char_cols <- baseline %>%
  select(-SUBJID,-AGE,-ARMCD) %>%
  names()

# generate list of dfs containing the results for the table
traits <- lapply(char_cols, cat_compare)

# bind everything into the table
groups_baseline <- bind_rows(age_tibble, traits)

knitr::kable(groups_baseline, format = "markdown")

```

The categorical variables for each arm were compared using a Fisher's exact test. The p-values are all greater than $\alpha$ = 0.05, except for stage. In other words, the groups of subjects in each arm are similar when looking at all of the baseline variables except stage. We have evidence that the composition of subjects at each stage of their sarcoma differs across arms. This may have an important impact on the outcome of this study if one considers stage to influence the likelihood of positive response to treatment.

**Criteria**:create a dataset named new_base after cleaning data to meet the criteria (delet 5 rows by filtering BECOG)
```{r}
#BECOG of 0 or 1
table(baseline$BECOG)#  the number of 0 is 365, the number of 1 is 270, the number of 2 is 4
sum(is.na(baseline$BECOG))# 1 NA in BECIG of Baseline_Information
# remove the 4(2) and 1(NA) from the BECOG of Baseline_Infor
#remove the BECOG which didn't meet the criteria
new_base <- baseline%>%
  filter(BECOG == 1|BECOG ==0)

```

**new_base Sheet** no NAs in ARMCD of the 635 patients (640-5(BECOG)=635)
```{r}
# new_base Sheet 
length(unique(new_base$ARMCD))
sum(is.na(new_base$ARMCD))
table0 <- data.frame(table(new_base$ARMCD))
table0 #317 patients in A, 323 patients in B, 640 in total

```


## Response Rate Comparison

### Last Response
**last_response Sheet**: 596 parients in last_response
No NAs in RSORRES of last_response
```{r}
# last_response Sheet
head(last_response)
dim(last_response) # 596 parients in last_response

# RSCAT
# check the total levels of RSCAT of last_response
length(unique(last_response$RSCAT))
# check if there is NAs in RSCAT of last_response
sum(is.na(last_response$RSCAT))# No Na in RSCAT

# RSORRES
# check the total levels of RSORRES of last_response
length(unique(last_response$RSORRES))
length(last_response$RSORRES) #Notice only 596 patients has the RSORRES in last_response
# check if there are NAs in RSORRES of last_response
sum(is.na(last_response$RSORRES)) # No NAs in RSORRES of last_response

# Create table1 showing frequency of each levels occurrence in RSORRES of last_response
table1 <- data.frame(table(last_response$RSORRES))
table1

```

inner_join Baseline_Information and last_response: ????check the outcome of inner_join 635-596-594
```{r}
# check the dims of both dataset 
dim(new_base)
dim(last_response)

#inner_join
Base_Last <- inner_join(last_response, new_base, by = "SUBJID")
dim(Base_Last)


#create data set inner_last only containing ARMCD,RSORRES
inner_last <- Base_Last %>%
  select(ARMCD,RSORRES)
unique(inner_last$RSORRES)
sum(is.na(inner_last$RSORRES))
```

last_response: Method_2by2
```{r}
# RSORRES: Response assessment, PD: Progressive disease, PR = Partial, response, CR = Complete response, SD = stable disease, NE = Inevaluable
inner_last%>%
  mutate(result =recode(RSORRES,"PD"="No Response", "PR"="Partial response", "SD" = "No Response", "CR"="Complete response", "NE" ="No Response")) -> inner_last_2by2

sampProps_last <- prop.table(table(inner_last_2by2$ARMCD, inner_last_2by2$result),1)
sampProps_last
table(inner_last_2by2$ARMCD, inner_last_2by2$result)
# sample difference
sampDiff_last <- sampProps_last[2,2]-sampProps_last[1,2]
sampDiff_last 

# Method1: chisq test
chisq.test(table(inner_last_2by2$ARMCD, inner_last_2by2$result), correct =F)

# Method2: relocation sampling
random.allo <-function(reps, group, outcome){
  propDiff <-rep(0,reps)
  for(i in 1:reps){
    randomGroup <-sample(group, replace=FALSE)
    propTable <-prop.table(table(randomGroup,outcome),1)
    propDiff[i] <- propTable[2,2]-propTable[1,2]
  }
  propDiff
}

sampleDiff_last <-random.allo(5000, inner_last_2by2$ARMCD, inner_last_2by2$result)
1-(sum(sampDiff_last< -(abs(sampleDiff_last))|sampDiff_last> abs(sampleDiff_last)))/5000
```

For the last response rate, at the significance level of 0.05, the response rate in treatment A is not significantly different with treatment B, with the p value of 0.85. 


### Best Response

**best_response Sheet**: 48  Nas of 640 patients in  BESTRESP of best_response 
only 592 patients could be used (640-48 = 592)
```{r}
# best_response Sheet
head(best_response)
dim(best_response) # 640 patients in best_response

# check the total levels of BESTRESP of best_response
length(unique(best_response$BESTRESP))
length(best_response$BESTRESP) #640 patients have the BESTRESP in best_response

# check if there are NAs in BESTRESP of last_response
sum(is.na(best_response$BESTRESP))# There are 48 NAs in BESTRESP of best_response

# Create table1 showing frequency of each levels occurrence in BESTRESPof last_response
table1 <- data.frame(table(best_response$BESTRESP))
table1
```

inner_join Baseline_Information and best_response
```{r}
# check the dims of both dataset
dim(baseline)
dim(best_response)

#inner_join
Base_Best <- inner_join(new_base, best_response, by = "SUBJID")
dim(Base_Best)

#create data set inner_last only containing ARMCD, BESTRESP
inner_best <- Base_Best %>%
  select(ARMCD,BESTRESP)

unique(inner_best$BESTRESP)
sum(is.na(inner_best$BESTRESP))# 45 NAs in inner_best, need I drop them to calculate the responce rate? I drop here

#omit the rows with NAs in inner_best to create the dataset inner_best_nonna
inner_best_nonna <- inner_best %>%
na.omit()
dim(inner_best)
dim(inner_best_nonna)
```



best_response: Method_2by2
```{r}
# RSORRES: Response assessment, PD: Progressive disease, PR = Partial, response, CR = Complete response, SD = stable disease, NE = Inevaluable
inner_best_nonna%>%
  mutate(result =recode(BESTRESP,"PD"="No Response", "PR"="Partial response", "SD" = "No Response", "CR"="Complete response", "NE" ="Inevaluable")) -> inner_best_nonna_2by2
dim(inner_best_nonna_2by2)

sampProps_best <- prop.table(table(inner_best_nonna_2by2$ARMCD,  inner_best_nonna_2by2$result),1)
sampProps_best 
table(inner_best_nonna_2by2$ARMCD,inner_best_nonna_2by2$result)
# sample difference
sampDiff_best <- sampProps_best[2,2]-sampProps_best[1,2]
sampDiff_best


# Method1: chisq test
chisq.test(table(inner_best_nonna_2by2$ARMCD, inner_best_nonna_2by2$result), correct = F)$p.value


# Method2: relocation sampling
random.allo <-function(reps, group, outcome){
  propDiff <-rep(0,reps)
  for(i in 1:reps){
    randomGroup <-sample(group, replace=FALSE)
    propTable <-prop.table(table(randomGroup,outcome),1)
    propDiff[i] <- propTable[2,2]-propTable[1,2]
  }
  propDiff
}

sampleDiff_best <-random.allo(5000, inner_best_nonna_2by2$ARMCD, inner_best_nonna_2by2$result)
1-(sum(sampDiff_best< -(abs(sampleDiff_best))|sampDiff_best> abs(sampleDiff_best)))/5000
```

For the best response rate, at the significance level of 0.05, the response rate in treatment A is significanly different with treatment B, with the p value of 0.0038. The response rate of A is higher than treatment B. 

```{r}
library(dplyr)
library(ggplot2)

# 1) recategorize：CR / PR / No Response
plot_df <- inner_best_nonna_2by2 %>%
  mutate(result3 = case_when(
    grepl("Complete", result, ignore.case = TRUE) ~ "Complete Response",
    grepl("Partial",  result, ignore.case = TRUE) ~ "Partial Response",
    TRUE ~ "No Response"   # SD/PD/NE/NR -> No Response
  )) %>%
  mutate(
    result3 = factor(result3,
                     levels = c("Complete Response","Partial Response","No Response")),
    ARMCD   = factor(ARMCD, labels = c("Combination Treatment Group",
                                       "Doxorubicin Alone Group"))
  ) %>%
  count(ARMCD, result3, name = "n") %>%
  group_by(ARMCD) %>%
  mutate(pct = n / sum(n))

# 2) drawing
ggplot(plot_df, aes(x = ARMCD, y = pct, fill = result3)) +
  geom_col(width = 0.7) +
  geom_text(aes(label = scales::percent(pct, accuracy = 0.01)),
            position = position_stack(vjust = 0.5), size = 3) +
  scale_y_continuous(labels = scales::percent,
                     limits = c(0,1), expand = c(0,0)) +
  scale_fill_manual(values = c("#00E6FF", "#2BBBD8", "#FFD23F"), name = NULL) +
  labs(x = "Treatment Group", y = "Response Rate",
       title = "Figure 1 Response Rates by Treatment Groups (%)") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom",
        plot.title = element_text(hjust = 0.5, face = "bold.italic"))
```


## Summarise Adverse Event Data

```{r}
# Merge adverse_event and baseline.info dataset. (Select the variables in the dictionary first)
adverse_event <- adverse_event %>%
  inner_join(baseline) %>%
  filter(BECOG %in% c(0,1)) %>%
  arrange(SUBJID) %>%
  select(SUBJID, AESOC, AESER, AEOUT, AETOXGR, STAGE, ARMCD)

```


```{r}
aesoc <- adverse_event %>%
  group_by(AESOC) %>%
  summarise(n = n()) %>%
  arrange(desc(n))
# The number of the unique AESOCs.
nrow(aesoc)
# The proportion of top 10 AESOCs.
prop_top10_aesoc <- sum(aesoc[1:10,2]/nrow(adverse_event))
round(prop_top10_aesoc, 4)
# The names of top 10 AESOCs.
top10_aesoc <- aesoc[1:10,1]
top10_aesoc
```

* The top 10 among 24 different system organ classes constitute more than 90% of the whole adverse event record size. So it is proper to filter those 10 in the following analysis.


```{r}
ae_file <- adverse_event %>%
  select(ARMCD, AESOC, AESER, AEOUT, AETOXGR) %>%
  # AETOXGR is originally a numeric and is supposed to be a factor.
  mutate(AETOXGR = as.factor(AETOXGR)) %>%
  # Select the top 10 AESOC and order the result.
  filter(AESOC %in% unlist(top10_aesoc)) %>%
  mutate(AESOC = factor(AESOC, levels = unlist(top10_aesoc)))
ae_file
```

## Table. AE Profile Comparison between Arm A and B.
```{r, results = 'asis'}
# All the four characteristics are categorial variables. Test is chisq by default.
summary(tableby(ARMCD ~ AESER + AEOUT + AETOXGR + AESOC, data = ae_file))
```




